Supplementary Material

### 2025-05-28 manuscript/manuscript.qmd
additions on runner up clinical trial candidates for emulation

While most trial eligibility criteria could be operationalized in an automated fashion, our final criterion was an assessment of emulation feasibility and clinical relevance. This criterion involved extensive human review. The fundamental points considered in this step include an initial feasibility assessment of the data fitness,[@rivera2022friends] including assessment of whether critical eligibility criteria (e.g., biomarker status) including prognostic factors (e.g., ECOG performance score) are measurable, and whether preliminary study size based on number of patients observed in the data with the combination of treatment regimen and line of treatment is reasonable. Lastly, trial candidates were ranked and shortlisted into primary and runner-up candidates based on their clinical and regulatory relevance.

A list of tentative, shortlisted primary candidates is presented in @tbl-rcts and the corresponding selection process is illustrated in the CONSORT diagrams (Supplementary Figures 1-4). Naturally, the majority of trials cover advanced (locally advanced, inoperable, recurrent/progressive disease) or metastatic cancer populations because a large proportion of drug development efforts have recently focused on these settings. A key objective that we aim to explore with the selected trials is to achieve a better understanding how different disease settings (early, late), line settings (\[neo\]adjuvant, first line, advanced lines of therapy), therapy protocols (monotherapy, combination therapy) and clinical characteristics (all-comer versus complex genetic or immunological signatures) can be emulated using RWD. If more thorough feasibility assessments after applying all relevant eligibility criteria suggest that the threat of bias from mis-measurement of key study parameters or residual confounding remains high, or that the study size is not sufficient, then runner-up candidate trials will be considered instead.

>>>>

While most trial eligibility criteria could be operationalized in an automated fashion, our final criterion was an assessment of emulation feasibility and clinical relevance. This criterion involved extensive human review. The fundamental points considered in this step include an initial feasibility assessment of the data fitness,[@rivera2022friends] involving an assessment of whether critical eligibility criteria (e.g., biomarker status) including prognostic factors (e.g., ECOG performance score) are measurable, and whether preliminary study size based on a rough estimation of the number of patients observed in the data with the combination of treatment regimen and line of treatment is reasonable for shortlisting the trial. Lastly, trial candidates were ranked and shortlisted into primary and runner-up candidates based on their clinical and regulatory relevance.

A list of tentative, shortlisted primary candidates is presented in @tbl-rcts. The corresponding selection process is illustrated in the CONSORT diagrams (**Supplementary Figures 1-4**) and runner up candidate trials are listed in **Supplementary Table 1**. Naturally, the majority of shortliested trials cover advanced (locally advanced, inoperable, recurrent/progressive disease) or metastatic cancer populations because a large proportion of drug development efforts have recently focused on these settings. A key objective that we aim to explore with the selected trials is to achieve a better understanding how different disease settings (early, late), line settings (\[neo\]adjuvant, first line, advanced lines of therapy), therapy protocols (monotherapy, combination therapy) and clinical characteristics (all-comer versus complex genetic or immunological signatures) can be emulated using RWD. If more thorough feasibility assessments after applying all relevant eligibility criteria suggest that the threat of bias from mis-measurement of key study parameters or residual confounding remains high, or that the study size is not sufficient, then runner-up candidate trials will be considered instead.

### 2025-05-28 manuscript/supplement.qmd
additions on runner up clinical trial candidates for emulation

library(here)
library(knitr)

@@

# Supplementary Figures

@@

>>>>

library(here)
library(knitr)
library(readxl)
library(gt)
library(dplyr)

@@

# Supplementary Figures and Tables

@@

```{r}
#| label: tbl-runner-ups
#| tbl-cap: "Oncology trial emulation potential runner up candidates."
#| tbl-cap-location: top

# read excel blueprint
runner_ups <- read_excel(
  path = here("tables", "helper_tables.xlsx"),
  sheet = "runner_ups"
  )

# convert to gt table
runner_ups_gt <- runner_ups |> 
  group_by(cancer) |> 
  gt() |> 
  cols_label(
    nctid = md("**NCTID**"),
    acronym = md("**Acronym**"),
    setting = md("**Clinical setting**"),
    line = md("**Line of therapy**"),
    exposures = md("**Treatment comparison**")
    ) |> 
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_row_groups()
    )

# save image to display within manuscript document
suppressMessages(runner_ups_gt |> 
  gtsave(
    filename = here("tables", "STable_runner_ups.png"),
    expand = 10, # amount of whitespace,
    zoom = 5 # higher resolution
    ))

include_graphics(path = here("tables", "STable_runner_ups.png"))
```
